---
title: "Lab6"
author: {Heather Childers}
date: "2023-03-01"
output: html_document
---

## Case Study: Eel Distribution Modeling

This week's lab follows a project modeling the eel species Anguilla australis described by Elith et al. (2008). There are two data sets for this lab. You'll use one for training and evaluating your model, and you'll use your model to make predictions predictions on the other. Then you'll compare your model's performance to the model used by Elith et al.

```{r, message = FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(ranger)
library(spData)
library(ggpmisc)
library(vip)
library(xgboost)
```

## Data

Grab the training data sets (eel.model.data.csv, eel.eval.data.csv) from github here: <https://github.com/MaRo406/eds-232-machine-learning/blob/main/data>

```{r, message=FALSE, warning=FALSE}
eel_dat <- read_csv("eel.model.data.csv") %>% 
  mutate(Angaus = as.factor(Angaus))
eval <- read_csv("eel.eval.data.csv") %>% 
  mutate(Angaus_obs = as.factor(Angaus_obs))
```

### Split and Resample

Split the model data (eel.model.data.csv) into a training and test set, stratified by outcome score (Angaus). Use 10-fold CV to resample the training set.

```{r}
set.seed(456)
#Split the model Data
eel_split <- initial_split(eel_dat)
eel_train <- training(eel_split)
eel_test <- testing(eel_split)

#Create the 10 fold cross validation
folds <- vfold_cv(eel_train, v = 10)
```

### Preprocess

Create a recipe to prepare your data for the XGBoost model

```{r}
eel_recipe = recipe(Angaus ~ ., data = eel_train) %>% #create model recipe
  step_dummy(all_nominal_predictors()) %>% #create dummy variables from all factors
  step_normalize(all_numeric_predictors()) #normalize all numeric predictors

```

## Tuning XGBoost

### Tune Learning Rate

Following the XGBoost tuning strategy outlined in lecture, first we conduct tuning on just the learning rate parameter:

1.  Create a model specification using {xgboost} for the estimation
    -   Only specify one parameter to tune()

```{r}
xgb_model = boost_tree(learn_rate = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

xgb_workflow = workflow() %>% #create workflow
  add_model(xgb_model) %>% #add boosted trees model
  add_recipe(eel_recipe) #add recipe

xgb_workflow
```

1.  Set up a grid to tune your model by using a range of learning rate parameter values: expand.grid(learn_rate = seq(0.0001, 0.3, length.out = 30))

-   Use appropriate metrics argument(s) - Computational efficiency becomes a factor as models get more complex and data get larger. Record the time it takes to run. Do this for each tuning phase you run. You could use {tictoc} or Sys.time().

```{r}
expanded_grid <- expand.grid(learn_rate = seq(0.0001, 0.3, length.out = 30))

system.time({
xgb_cv_tune = xgb_workflow %>%
  tune_grid(resamples = folds, grid = expanded_grid) #use cross validation to tune learn_rate and trees parameter
})
```

3.  Show the performance of the best models and the estimates for the learning rate parameter values associated with each.

```{r}
collect_metrics(xgb_cv_tune) #get metrics from tuning cv to pick best model

autoplot(xgb_cv_tune) + #plot cv results for parameter tuning
  theme_bw()

xgb_best = show_best(xgb_cv_tune, n = 1, metric = "accuracy") #get metrics for best random forest model
xgb_best
```

### Tune Tree Parameters

1.  Create a new specification where you set the learning rate (which you already optimized) and tune the tree parameters.

```{r}

xgb_model2 = boost_tree(learn_rate = 0.1241966, 
                        trees = 3000, 
                        tree_depth = tune(), 
                        min_n = tune(), 
                        loss_reduction = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")
xgb_workflow2 = workflow() %>% #create workflow
  add_model(xgb_model2) %>% #add boosted trees model
  add_recipe(eel_recipe) #add recipe
```

1.  Set up a tuning grid. This time use grid_latin_hypercube() to get a representative sampling of the parameter space

NOTE : if no grid is specified, tune_grid will use grid_latin_hypercube()

```{r}
xgb2_cv_tune = xgb_workflow2 %>%
  tune_grid(resamples = folds)
```

1.  Show the performance of the best models and the estimates for the tree parameter values associated with each.

```{r}
xgb_best2 = show_best(xgb2_cv_tune, n = 1, metric = "accuracy") #get metrics for best random forest model
xgb_best2
```

### Tune Stochastic Parameters

1.  Create a new specification where you set the learning rate and tree parameters (which you already optimized) and tune the stochastic parameters.

    ```{r}

    xgb_model3 = boost_tree(learn_rate = 0.1241966, 
                            trees = 3000, 
                            tree_depth = 4, 
                            min_n = 3, 
                            loss_reduction = 0.00048,
                            mtry = tune(),
                            sample_size = tune()) %>% 
      set_engine("xgboost") %>% 
      set_mode("classification")



    xgb_workflow3 = workflow() %>% #create workflow
      add_model(xgb_model3) %>% #add boosted trees model
      add_recipe(eel_recipe) #add recipe

    xgb <- xgb_model3 %>% 
      extract_parameter_set_dials()

    ```

2.  Set up a tuning grid. Use grid_latin_hypercube() again.

```{r}

tune <- xgb_workflow3 %>%  
  tune_grid(resamples = folds,
            grid = grid_latin_hypercube(finalize(xgb, x = eel_train)),
                                        size = 10)

show_best(tune, n = 1, metric = "accuracy")
```

1.  Show the performance of the best models and the estimates for the tree parameter values associated with each.

## Finalize workflow and make final prediction

1.  How well did your model perform? What types of errors did it make?

The model performed fairly well, The accuracy for the model is 82.8%

## Fit your model the evaluation data and compare performance

1.  Now used your final model to predict on the other dataset (eval.data.csv)

```{r}
eel_recipe2 = recipe(Angaus_obs ~ ., data = eval) %>% #create model recipe
  step_dummy(all_nominal_predictors()) %>% #create dummy variables from all factors
  step_normalize(all_numeric_predictors()) #normalize all numeric predictors
```

1.  How does your model perform on this data?

```{r}
xgb_workflow4 = workflow() %>% #create workflow
  add_model(xgb_model3) %>% #add boosted trees model
  add_recipe(eel_recipe2) #add recipe


#final_fit <- xgb_workflow3 %>% fit(data = eval)

```
